{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-07-28 thesis/tm.tex
extend verifs table and run buncha small retry-set and abort-code verif tests

Up to three abort codes are considered depending on the testing options listed in the previous section.
All abort codes are simultaneously added to both lists,
unless already present on {\tt xabort\_codes\_ever},
in which case not added to {\tt xabort\_codes\_todo} to avoid duplicate work.
\begin{itemize}
	\item {\tt \_XABORT\_RETRY}:
		When a {\tt xbegin} preemption point is created ({\tt save\_setjmp()}),
		if {\tt -S} is not set,
		both {\tt xabort\_codes} lists are initialized to contain this code.
		This represents the possibility for a hardware transaction to fail
		for reasons outside of the programmer's control such as system interrupts or cache eviction.
		If {\tt -S} is set they are initialized to empty,
		representing either STM's policy of failing only when a true memory conflict arises,
		or a TSX user wrapping all their {\tt \_xbegin()}s in a retry loop such as in \cite{sigbovik-htm}.
	\item {\tt \_XABORT\_CONFLICT}:
		Whenever DPOR detects a memory conflict between two transitions ({\tt shimsham\_shm()}),
		if {\tt -A} is set,
		% for each of the two which are within a transaction, % no!!!!!!!
		%it finds the nearest ancestor {\tt xbegin} preemption point by the same thread
		if the later-executed transition is a transactional success path,
		it adds this code to both lists of the immediately preceding preemption point.
		Note that a transaction should suffer a conflict abort
		only when executed {\em after} a conflicting memory access
		to avoid circular causality
		(see {\tt 410user/progs/htm\_causality.c} for rationale).
		Should the transaction happen to be executed first,
		DPOR will first try to reorder it as normal,
		and then abort it as described if the conflict persists.
	\item {\tt \_XABORT\_EXPLICIT}:
		Whenever the program invokes {\tt \_xabort()} ({\tt sched\_update\_user\_\allowbreak{}state\_machine()}),
		if {\tt -A} is set,
		this code is added to both lists, bitwise-ored with the user-supplied argument code
		as specified in \cite{htm-gcc}.
\end{itemize}

@@

To solve this,
I extended Landslide
with an experimental feature called {\em retry set reduction},
so named after the analogous {\em sleep sets} from prior work
(\cite{partial-order-methods,dpor,optimal-dpor}, \sect{\ref{sec:landslide-sleepsets}}).
Whenever DPOR tags a new branch for exploration (\sect{\ref{sec:landslide-explore}}),
if either or both sides of a memory conflict were part of a transaction,
it records a {\em retry set},
i.e., a pair of {\tt xbegin} results executed by the conflicting threads.
% ah-- not compatible after all!
%in addition to tracking the set of abort codes possible at each preemption point.
Later, when traversing that new branch,
Landslide remembers those threads' expected {\tt xbegin} results
and refuses to test any others
(unless DPOR separately found them, too, in conflict),
thereby skipping over (i.e., pruning) reorderings of any other {\tt xbegin} results that would be independent.
The {\tt landslide} repository's {\tt retry-sets} branch implements this feature%
\footnote{The upcoming evaluation showed it too unstable, let alone not yet proved sound, to include in {\tt master}.}.

@@

\begin{table}[h]
	\begin{center}
		\footnotesize
		\begin{tabular}{cc|r|r}
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf params & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
			\hline
			\hline
			{\tt htm2}
			& 2,1 & \cpu{16.30}	& \ints{22}	\\
			& 2,2 & \cpu{69.80}	& \ints{1970}	\\
			& 2,3 &\cpu{3510.00}	& \ints{104914}	\\
			%& 2,4 &	\ETAdag{...}	& \ETAdag{...}	\\
			& 3,1 &\cpu{41.07}	& \ints{941}	\\
			& 3,2 &	\ETAdag{9d 19h}	& \ETAdag{5058149}	\\ % add 10h ok
			% [JOB 4] progress: 923822/5058149 brs (18.264030%), ETA 9d 8h 30m 38s (elapsed 9h 59m 55s)
			& 4,1 & \cpu{12699.65}	& \ints{344240}	\\
			\hline
			{\tt counter}
			& 2,1 & \cpu{5.40}	& \ints{10}	\\
			& 2,2 & \cpu{10.91}	& \ints{190}	\\
			& 2,3 & \cpu{134.21}	& \ints{3970}	\\
			& 2,4 & \cpu{3009.40}	& \ints{86950}	\\
			& 3,1 & \cpu{8.30}	& \ints{120}	\\
			& 3,2 & \cpu{2038.90}	& \ints{60606}	\\
			& 4,1 & \cpu{101.95}	& \ints{3006}	\\
			\hline
			{\tt swap}
			& 2,1 & \cpu{38.79}	& \ints{99}	\\
			& 2,2 & \cpu{10299.27}	& \ints{233396}	\\
			& 3,1 & \ETAdag{todo} & \ETAdag{todo}	\\
			\hline
			{\tt fig63}
			& 2,1 & \cpu{todo}	& \ints{todo}	\\
			& 2,2 & \cpu{todo}	& \ints{todo}	\\
			& 3,2 & \ETAdag{todo} & \ETAdag{todo}	\\
			\hline
			{\tt avl\_insert}
			& 2,1 & \cpu{491.98}	& \ints{15056}	\\
			\hline
			{\tt avl\_fixed}
			& 2,1 & \cpu{572.49} 	& \ints{20420} 	\\
			& 2,2 & \ETAdag{133y}	& \ETAdag{51484570} 	\\ % added 10hr but dosent matter
			% [JOB 19] progress: 408805/51484570 brs (0.794034%), ETA 132y 356d 8h 55m 16s (elapsed 9h 59m 19s)
			& 3,1 & \ETAdag{todo}	& \ETAdag{todo}	\\

			\hline
			{\tt map\_basic}
			& 2,1 & \ETAdag{todo} & \ETAdag{todo} \\
			\hline
			{\tt map\_basicer}
			& 2,1 & \cpu{36.17}& \ints{426}	\\ % TODO make sure this gets updated if you update the test
			& 2,2 & \ETAdag{todo} & \ETAdag{todo} \\
			& 3,1 & \ETAdag{todo} & \ETAdag{todo} \\
		\end{tabular}
	\end{center}
	\caption{Transactional tests verified (or not) by Landslide.
		Run with {\tt landslide -M -X}, i.e., retry aborts enabled, different xabort codes not distinguished.
		State space estimates measured after a timeout of 10 hours.
		}
	\label{tab:verifs}
\end{table}

>>>>

Up to three abort codes are considered depending on the testing options listed in the previous section.
All abort codes are simultaneously added to both lists,
unless already present on {\tt xabort\_codes\_ever},
in which case not added to {\tt xabort\_codes\_todo} to avoid duplicate work.
\begin{itemize}
	\item {\tt \_XABORT\_RETRY}:
		When a {\tt xbegin} preemption point is created ({\tt save\_setjmp()}),
		if {\tt -S} is not set,
		both {\tt xabort\_codes} lists are initialized to contain this code.
		This represents the possibility for a hardware transaction to fail
		for reasons outside of the programmer's control such as system interrupts or cache eviction.
		If {\tt -S} is set they are initialized to empty,
		representing either STM's policy of failing only when a true memory conflict arises,
		or a TSX user wrapping all their {\tt \_xbegin()}s in a retry loop such as in \cite{sigbovik-htm}.
	\item {\tt \_XABORT\_CONFLICT}:
		Whenever DPOR detects a memory conflict between two transitions ({\tt shimsham\_shm()}),
		if {\tt -A} is set,
		% for each of the two which are within a transaction, % no!!!!!!!
		%it finds the nearest ancestor {\tt xbegin} preemption point by the same thread
		if the later-executed transition is a transactional success path,
		it adds this code to both lists of the immediately preceding preemption point.
		Note that a transaction should suffer a conflict abort
		only when executed {\em after} a conflicting memory access
		to avoid circular causality
		(see {\tt 410user/progs/htm\_causality.c} for rationale).
		Should the transaction happen to be executed first,
		DPOR will first try to reorder it as normal,
		and then abort it as described if the conflict persists.
	\item {\tt \_XABORT\_EXPLICIT}:
		Whenever the program invokes {\tt \_xabort()} ({\tt sched\_update\_user\_\allowbreak{}state\_machine()}),
		if {\tt -A} is set,
		this code is added to both lists, bitwise-ored with the user-supplied argument code
		as specified in \cite{htm-gcc},
		and execution of the transactional path immediately stops.
	\item {\tt \_XABORT\_CAPACITY}:
		Whenever the program invokes a system call during a transaction,
		if {\tt -A} is set,
		this code is added to both lists,
		and execution of the transactional path immediately stops.
\end{itemize}

@@

To solve this,
I extended Landslide
with {\em retry set reduction},
an experimental feature named after the analogous {\em sleep sets} from prior work
(\cite{partial-order-methods,dpor,optimal-dpor}, \sect{\ref{sec:landslide-sleepsets}}).
Whenever DPOR tags a new branch for exploration (\sect{\ref{sec:landslide-explore}}),
if either or both sides of a memory conflict were part of a transaction,
it records a {\em retry set},
i.e., the pair of {\tt xbegin} results executed by the conflicting threads.
% ah-- not compatible after all!
%in addition to tracking the set of abort codes possible at each preemption point.
Later, when traversing that new branch,
Landslide remembers those threads' expected {\tt xbegin} results
and refuses to test any others
(unless DPOR separately found them, too, in conflict),
thereby skipping over (i.e., pruning) reorderings of any other {\tt xbegin} results that would be independent.
The {\tt landslide} repository's {\tt retry-sets} branch implements this feature%
\footnote{The upcoming evaluation showed it too unstable, let alone not yet proved sound, to include in {\tt master}.},
available with the {\tt -R} command-line option.

@@

% TODO
% discuss the diff reducion approaches

% Please note that retry sets are a theoretically-founded reduction that should apply to all programs,
% while distinguishing abort codes to suppress retries is potentially unsound,
% depending on the user's ability to manually verify the presence (and correctness) of retry loops in the code.

\begin{table}[h]
	\begin{center}
		\footnotesize
		\begin{tabular}{cc||r|r||r|r||r|r}
			& & \multicolumn{2}{c||}{\bf Retry aborts only}
			  & \multicolumn{2}{c||}{\bf Retry sets ({\tt -R})}
			  & \multicolumn{2}{c}{\bf Abort codes ({\tt -A -S})} \\
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size}
			  & \cpu{\bf cpu (s)} & \ints{\bf SS size}
			  & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf params
			  & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.})
			  & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.})
			  & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
			\hline
			\hline
			{\tt htm2}
			& 2,1 & \cpu{16.30}	& \ints{22}	& \cpu{19.82}	& \ints{17}	& \cpu{3.32}	& \ints{4} \\
			& 2,2 & \cpu{69.80}	& \ints{1970}	& \cpu{38.71}	& \ints{906}	& \cpu{26.59}	& \ints{288} \\
			& 2,3 &\cpu{3510.00}	& \ints{104914}	& \cpu{}	& \ints{}	& \cpu{}	& \ints{} \\
			%& 2,4 &	\ETAdag{...}	& \ETAdag{...}	\\
			& 3,1 &\cpu{41.07}	& \ints{941}	& \cpu{29.38}	& \ints{553}	& \cpu{23.03}	& \ints{162} \\
			& 3,2 &	\ETAdag{9d 19h}	& \ETAdag{5058149}	\\ % add 10h ok
			% [JOB 4] progress: 923822/5058149 brs (18.264030%), ETA 9d 8h 30m 38s (elapsed 9h 59m 55s)
			& 4,1 & \cpu{12699.65}	& \ints{344240}	& \cpu{}	& \ints{}	& \cpu{}	& \ints{} \\
			\hline
			{\tt counter}
			& 2,1 & \cpu{5.40}	& \ints{10}	& \cpu{6.70}	& \ints{8}	& \cpu{3.22}	& \ints{4} \\
			& 2,2 & \cpu{10.91}	& \ints{190}	& \cpu{14.91}	& \ints{288}	& \cpu{7.90}	& \ints{48} \\
			& 2,3 & \cpu{134.21}	& \ints{3970}	& \cpu{379.04}	& \ints{11512}	& \cpu{35.00}	& \ints{904} \\
			& 2,4 & \cpu{3009.40}	& \ints{86950}	& \cpu{}	& \ints{}	& \cpu{}	& \ints{} \\
			& 3,1 & \cpu{8.30}	& \ints{120}	& \cpu{8.90}	& \ints{74}	& \cpu{7.55}	& \ints{40} \\
			& 3,2 & \cpu{2038.90}	& \ints{60606}	& \cpu{}	& \ints{}	& \cpu{}	& \ints{} \\
			& 4,1 & \cpu{101.95}	& \ints{3006}	& \cpu{99.22}	& \ints{2854}	& \cpu{32.95}	& \ints{848} \\
			\hline
			{\tt swap}
			& 2,1 & \cpu{38.79}	& \ints{99}	& \cpu{35.80}	& \ints{84}	& \cpu{3.25}	& \ints{4} \\
			& 2,2 & \cpu{10299.27}	& \ints{233396}	& \cpu{}	& \ints{}	& \cpu{}	& \ints{} \\
			& 3,1 & \ETAdag{todo}	& \ETAdag{todo}	& \cpu{}	& \ints{}	& \cpu{}	& \ints{} \\
			\hline
			{\tt fig63}
			& 2,1 & \cpu{6.66}	& \ints{18}	& \cpu{6.50}	& \ints{6}	& \cpu{3.10}	& \ints{1} \\
			& 2,2 & \cpu{11.52}	& \ints{244}	& \cpu{10.74}	& \ints{192}	& \cpu{3.12}	& \ints{1} \\
			& 2,3 & \cpu{107.04}	& \ints{4428}	& \cpu{146.96}	& \ints{6151}	& \cpu{3.13}	& \ints{1} \\
			& 3,1 & \cpu{13.68}	& \ints{296}	& \cpu{7.29}	& \ints{21}	& \cpu{3.26}	& \ints{1} \\
			& 3,2 & \ETAdag{todo}	& \ETAdag{todo}	& \cpu{}	& \ints{}	& \cpu{3.35}	& \ints{1} \\
			& 4,1 & \cpu{220.27}	& \ints{7382}	& \cpu{11.40}	& \ints{157}	& \cpu{3.27}	& \ints{1} \\
			\hline
			{\tt avl\_insert}
			& 2,1 & \cpu{491.98}	& \ints{15056}	& \cpu{220.53}	& \ints{6332}	& \cpu{98.77}	& \ints{2776} \\
			% (2,2) - bug found on regular w/ 127 ints, on retry sets w/ 201 ints, and codes w/ 45 ints
			\hline
			{\tt avl\_fixed}
			& 2,1 & \cpu{572.49} 	& \ints{20420} 	& \cpu{271.11}	& \ints{9739}	& \cpu{99.16}	& \ints{2776} \\
			& 2,2 & \ETAdag{133y}	& \ETAdag{51484570} 	\\ % added 10hr but dosent matter
			% [JOB 19] progress: 408805/51484570 brs (0.794034%), ETA 132y 356d 8h 55m 16s (elapsed 9h 59m 19s)
			& 3,1 & \ETAdag{todo}	& \ETAdag{todo}	\\

			\hline
			{\tt map\_basic}
			& 2,1 & \ETAdag{todo} & \ETAdag{todo} \\
			\hline
			{\tt map\_basicer}
			& 2,1 & \cpu{36.17}& \ints{426}	\\ % TODO make sure this gets updated if you update the test
			& 2,2 & \ETAdag{todo} & \ETAdag{todo} \\
			& 3,1 & \ETAdag{todo} & \ETAdag{todo} \\
		\end{tabular}
	\end{center}
	\caption{Transactional tests verified (or not) by Landslide.
		Run with {\tt -M -X}, plus any additional reduction options listed.
		``Retry aborts only'' always tests every abort path without distinguishing among failure reasons;
		``Retry sets'' skips equivalent retry abort reorderings (\sect{\ref{sec:tm-retrysets}});
		``Abort codes'' suppresses retries and dynamically detects when to inject conflict aborts and so on
		(\sect{\ref{sec:txn-abort-codes}}).
		State space estimates measured after a timeout of 10 hours.
		}
	\label{tab:verifs}
\end{table}
